---
title: "Assignment 2: Coding Basics"
author: "Ayoung Kim"
output: pdf_document
geometry: margin=2.54cm
editor_options: 
  chunk_output_type: console
---

## OVERVIEW

This exercise accompanies the lessons in Environmental Data Analytics on coding basics.

## Directions

1.  Rename this file `<FirstLast>_A02_CodingBasics.Rmd` (replacing `<FirstLast>` with your first and last name).
2.  Change "Student Name" on line 3 (above) with your name.
3.  Work through the steps, **creating code and output** that fulfill each instruction.
4.  Be sure to **answer the questions** in this assignment document.
5.  When you have completed the assignment, **Knit** the text and code into a single PDF file.
6.  After Knitting, submit the completed exercise (PDF file) to Sakai.

## Basics, Part 1

1.  Generate a sequence of numbers from one to 30, increasing by threes. Assign this sequence a name.

2.  Compute the mean and median of this sequence.

3.  Ask R to determine whether the mean is greater than the median.

4.  Insert comments in your code to describe what you are doing.

```{r basics 1}
#1. Sequence 1 (I defined seq(1,30.3) as sequence 1)
seq(1,30,3)
sequence1<-seq(1,30,3)

#2. Mean and Median (I got mean and median of Sequence 1 that I defined above)

mean(sequence1)
median(sequence1)

#3. Mean > Median ? (I got 14.5 for both mean and median)
14.5>14.5


## Basics, Part 2

5.  Create a series of vectors, each with four components, consisting of (a) names of students, (b) test scores out of a total 100 points, and (c) whether or not they have passed the test (TRUE or FALSE) with a passing grade of 50.

6.  Label each vector with a comment on what type of vector it is.

7.  Combine each of the vectors into a data frame. Assign the data frame an informative name.

8.  Label the columns of your data frame with informative titles.
```

```{r basics 2}
#testscore

#Name of Student
student = c("Anne","Marie","Jack","Wilson")

#Test score
testscore = c(45,70,80,90)

#Pass or Fail
pass<-(test_score>50)

#Test score/student name/pass
test_score<-testscore
student_name <- student
pass<-pass

#Data frame for Test Score
class(test_score)
df_test_score <- as.data.frame(test_score)
df_test_score

#Data frame for Student Names
class(student_name)
df_student_name <-as.data.frame(student_name)
df_student_name

#Data frame for 'Pass or Fail'
class(pass)
df_pass <-as.data.frame(pass)
df_pass

#Adding Columns 
df<-cbind(df_student_name,df_test_score,df_pass)
class(df)
df
names(df)<-"Test Scores of Students"

#Name
names(df_student_name)<-"Student"
names(df_test_score)<-"Score"
names(df_pass)<-"Pass"


9.  QUESTION: How is this data frame different from a matrix?

#Answer: with a data frame, I can combine data sets and show in a one chart. 

10. Create a function with an if/else statement. Your function should take a **vector** of test scores and print (not return) whether a given test score is a passing grade of 50 or above (TRUE or FALSE). You will need to choose either the `if` and `else` statements or the `ifelse` statement. 

#Function for Marie, Jack, and Wilson
df_test_score<-x
if(x>50){
  "Pass"
}
#Function for Anne
df_test_score<-x
if(x<50){
  'fail'
}


11. Apply your function to the vector with test scores that you created in number 5.

```{r basics 3}
#Test score for Anne
df_test_score <-45
if (df_test_score<50){
  'Fail'
}

#Test score for Marie
df_test_score<-70
if(df_test_score>50){
  'pass'
}
#Test score for Jack
df_test_score <-80
if(df_test_score>50){
  'pass'
}
#Test score for Wilson
df_test_score<-90
if(df_test_score>50){
  'pass'
}

```

12. QUESTION: Which option of `if` and `else` vs. `ifelse` worked? Why?

> Answer: The 'if' option worked for me. I used different function for Anne's score.I put x<50 for Anne's score and x>50 for the scores of Marie, Jack, and Wilson.I thought, for Anne's score, the combination of 'if' and 'else' would work, but it didn't. so I decided to put x<50 for Anne's case. 
